\thetitle

### 2021-05-26 writing/discussion.tex
commenting out soem margin notes + removing duplicate threat to validity listing

The validity of these conclusions is threatened in two fundamentally distinct
ways. The first sets of threats have been mentioned before: (i) the
representativeness of benchmarks; (ii) the relation between mutations and real
programming mistakes; (iii) the definition of interesting debugging scenarios;
and (iv) the sampling strategy.  While the experimental setup attempts to
mitigate against these threats, the reader must definitely take into account
these limitations of the experiment and must understand that the collected data
not support any further generalization.

>>>>

The validity of these conclusions is threatened in two fundamentally distinct
ways. The first sets of threats have been mentioned before: (i) the
representativeness of benchmarks; (ii) the relation between mutations and real
programming mistakes; and (iii) the sampling strategy.  While the experimental setup attempts to
mitigate against these threats, the reader must definitely take into account
these limitations of the experiment and must understand that the collected data
not support any further generalization.

### 2021-05-26 writing/landscape.tex
commenting out soem margin notes + removing duplicate threat to validity listing

As this analysis implies, if a return-type check fails, the problem is
that the untyped module supplied a function, here {\tt
pack-lib}, that is not a  match for the type ascribed by the typed module. 
Hence either the type at the boundary between the two modules is wrong or, 
if the programmer trusts the type, the untyped module is at fault.
If an argument-type check fails, responsibility  lies with
{\tt client}. After all either the type it ascribes to the argument is
wrong or the argument it produced clashes with the type. Due to 
proxies, {\it Natural\/} can easily track the 
boundary, type and responsible parties that correspond to each check. 
Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary between \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypto-pack-lib}, respectively, for the two {\tt define}s in {\tt client}.

@@

The following table summarizes the illustration. Each cell describes the result
of evaluating the column's definition (in {\tt client}) under the row's
semantics.
%% -----------------------------------------------------------------------------
\begin{center}
  \begin{tabular}{l|l@{\qquad}l}
                  & {\tt public-data}              & {\tt secret-data}                                 \\
\hline %% ----------------------------------------------------------------------------------------
{\it Natural\/}   & error, blaming the boundary between & error, blaming the boundary between          \\
                  & {\tt pack-lib} and {\tt typed-pack-lib} & {\tt pack-lib} and {\tt crypto-pack-lib} \\
\hline
{\it Transient\/} & no error                       & error, blaming the boundaries between            \\
                  &                                & {\tt pack-lib}  and {\tt typed-pack-lib}, and \\
                  &                                & {\tt  pack-lib} and {\tt crypto-pack-lib}     \\
\hline
 {\it Erasure\/}   & no error*                      & no error*                                         \\
                  &   \multicolumn{2}{|c}{*but {\em Erasure} does signal an error on list access}
\end{tabular}
\end{center}
%% -----------------------------------------------------------------------------

>>>>

As this analysis implies, if a return-type check fails, the problem is
that the untyped module supplied a function, here {\tt
pack-lib}, that is not a  match for the type ascribed by the typed module. 
Hence either the type at the boundary between the two modules is wrong or, 
if the programmer trusts the type, the untyped module is at fault.
If an argument-type check fails, responsibility  lies with
{\tt client}. After all, either the type it ascribes to the argument is
wrong or the argument it produced clashes with the type. Due to 
proxies, {\it Natural\/} can easily track the 
boundary, type and responsible parties that correspond to each check. 
Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary between \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypto-pack-lib}, respectively, for the two {\tt define}s in {\tt client}.

@@

The following table summarizes the illustration. Each cell describes the result
of evaluating the column's definition (in {\tt client}) under the row's
semantics.
%% -----------------------------------------------------------------------------
\begin{center}
  \begin{tabular}{l|l@{\qquad}l}
                  & {\tt public-data}              & {\tt secret-data}                                 \\
\hline %% ----------------------------------------------------------------------------------------
{\it Natural\/}   & error, blaming the boundary between & error, blaming the boundary between          \\
                  & {\tt pack-lib} and {\tt typed-pack-lib} & {\tt pack-lib} and {\tt crypto-pack-lib} \\
\hline
{\it Transient\/} & no error                       & error, blaming the boundaries between            \\
                  &                                & {\tt pack-lib}  and {\tt typed-pack-lib}/ \\
                  &                                & {\tt  pack-lib} and {\tt crypto-pack-lib}     \\
\hline
 {\it Erasure\/}   & no error*                      & no error*                                         \\
                  &   \multicolumn{2}{|c}{*but {\em Erasure} does signal an error on list access}
\end{tabular}
\end{center}
%% -----------------------------------------------------------------------------

### 2021-05-26 writing/rational-natural.tex
commenting out soem margin notes + removing duplicate threat to validity listing

\mf{Technically a Natural blame trail could be a Natural exception trail. Hmph? }
\mf{I added a sentence below this definition and the next to address this issue.}

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\blame{\system, \conf_i}\}$      & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$  & otherwise \\ %
\end{tabular} \right. %
\]
where $\blame{\system, \conf}$ denotes the module (of $\system$) that $\conf$ blames,
and $\exception{\system, \conf}$ denotes the first untyped location in the stacktrace produced by $\conf$.
\end{quote}
Note how this definition interprets ``blame information'' broadly, as
any information from a failed run-time check in a blame-oriented semantics.

This definition rests on the two options for $\conf_{i+1} \setminus \conf_i$.
The first part reflects the rational programmer's use of blame to extend the
trail. However, blame may not be available in some scenarios. The second part
accounts for those scenarios that produce errors from the runtime system before
any impedance mismatch can be detected. For example, the fully untyped
configuration can only produce such an error, e.g. from {\tt +} receiving a
string. Exceptions from the runtime do not carry blame, so
the rational programmer proceeds using the accompanying stacktrace instead.

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based entirely on exceptions.
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$.
\end{quote}
Using Natural exception trails it becomes possible to factor out
``blame information'' in the narrow sense from the broad one of the
above definition.

>>>>

%\mf{Technically a Natural blame trail could be a Natural exception trail. Hmph? }
%\mf{I added a sentence below this definition and the next to address this issue.}
%\cd{I revised a bit.}

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\blame{\system, \conf_i}\}$      & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$  & otherwise \\ %
\end{tabular} \right. %
\]
where $\blame{\system, \conf}$ denotes the module (of $\system$) that $\conf$ blames,
and $\exception{\system, \conf}$ denotes the first untyped location in the stacktrace produced by $\conf$.
\end{quote}
Note how, in the absence of ``blame information'' in the narrow sense, 
this definition interprets ``blame information'' broadly, as in
any information from a failing run-time check.  
In particular, the definition rests on the two options for $\conf_{i+1} \setminus \conf_i$.
The first part reflects the rational programmer's use of blame to extend the
trail. However, blame may not be available in some scenarios. The second part
accounts for those scenarios that produce errors from the runtime system before
any impedance mismatch can be detected. For example, the fully untyped
configuration can only produce such an error, e.g. from {\tt +} receiving a
string. Exceptions from the runtime do not carry blame, so
the rational programmer proceeds using the accompanying stacktrace instead.

@@

To account for this situation, a new mode of the Natural rational
programmer follows a migration process based entirely on exceptions.
\begin{quote}
\it A {\em Natural exception trail\/} is a sequence of scenarios $\conf_0,...\conf_n$ of a
program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and $\conf_{i+1} \setminus \conf_i = \{\exception{\system, \conf_i}\}$.
\end{quote}
Using Natural exception trails, it becomes possible to factor out
``blame information'' in the narrow sense from the broad one of the
above definition.
